---
title: 'EHSC P8321: Lab & Homework #1'
author: 'Lizbeth Gomez (UNI: LG2982'
date: 'Due: April 6, 2020 by 3:59 pm'
output:
  word_document: default
  pdf_document: default
---

## PART I. Setting up RStudio & Reading Datasets in
### The first task we want to do after creating up our R Markdown document is to install and load any necessary packages.
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# install.packages("tidyverse")
library(tidyverse)

setwd("~/Desktop/CUMC-MPH/Spring 2020/INTRO TO DATA SCI/Session_1")
```


### Create new dataframes in RStudio with the COVID-19 time series datasets.
```{r loaddata}
confirmed = read_csv("./data/time_series_covid19_confirmed_global.csv")
deaths = read_csv("./data/time_series_covid19_deaths_global.csv")
recovered = read_csv("./data/time_series_covid19_recovered_global.csv")
```



## PART II. Descriptive Statistics
### What if we wanted to know the total number of confirmed cases, deaths, and those recovered from COVID-19 of a country? Let's generate some summary statistics for Iran.
```{r iran_statistics}
# Let's do some data tidying. First we want to convert the dataset from wide-to-long format using the pivot_longer function. We can use the janitor::clean_names() function to change the variable names into RStudio-friendly names. Next, we want to select only the variables of interest. Last, we can filter the dataset for cases in Iran only.
iran_confirmed = confirmed %>%
  pivot_longer(
    "1/22/20":"3/7/20",
    names_to = "date",
    values_to = "cases") %>%
  janitor::clean_names() %>% 
  select(country_region, lat, long, date, cases) %>% 
  filter (country_region == "Iran") 
iran_totalcases = sum(iran_confirmed$cases)

# We can follow the same tidying method for the deaths and recovered datasets.
iran_deaths = deaths %>%
  pivot_longer(
    "1/22/20":"3/7/20",
    names_to = "date",
    values_to = "deaths") %>%
  janitor::clean_names() %>% 
  select(country_region, lat, long, date, deaths) %>% 
  filter (country_region == "Iran")
iran_totaldeaths = sum(iran_deaths$deaths)

iran_recovered = recovered %>%
  pivot_longer(
    "1/22/20":"3/7/20",
    names_to = "date",
    values_to = "recovered") %>%
  janitor::clean_names() %>% 
  select(country_region, lat, long, date, recovered) %>% 
  filter (country_region == "Iran")
iran_totalrecovered = sum(iran_recovered$recovered)
```
To date, Iran has `r iran_totalcases` total confirmed cases, `r iran_totaldeaths` total deaths, and `r iran_totalrecovered` total recovered individuals.


### Your turn: Generate the total number of confirmed cases, deaths, and those recovered from COVID-19 in Ireland. Using inline R, include a summary sentence about Ireland's statistics.
```{r ireland_statistics}

```




## PART III. Creating Visualizations
### Let's create a scatterplot with confirmed case counts to date of the three largest Chinese cities by population (Shanghai, Beijing, and Chongqing).
```{r scatterplot_china}
china_confirmed = confirmed %>%
  pivot_longer(
    "1/22/20":"3/7/20",
    names_to = "date",
    values_to = "cases") %>%
  janitor::clean_names() %>% 
  select(province_state, country_region, lat, long, date, cases) %>% 
  rename ("city" = province_state) %>% 
  filter (country_region == "China", city == c("Shanghai", "Beijing", "Chongqing")) %>% 
  mutate(date = as.Date(date, format = "%m/%d/%y"))

china_confirmed %>% 
  ggplot(aes(x = date, y = cases, color = city)) + 
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ city) +
  labs(title = "Confirmed COVID-19 Cases, 1/22/20-3/7/20",
       x = "Date",
       y = "Number of Confirmed Cases")
```


### Your turn: create a scatterplot of Italy's confirmed case counts to date.
```{r scatterplot_italy}

```


## What if we wanted to know if the distribution of the number of deaths from COVID-19 differed across the three main cities in China? Let's create a histogram of China's deaths from COVID-19 to date:
```{r histogram_china}
china_deaths = deaths %>%
  pivot_longer(
    "1/22/20":"3/7/20",
    names_to = "date",
    values_to = "deaths") %>%
  janitor::clean_names() %>% 
  select(province_state, country_region, lat, long, date, deaths) %>% 
  rename ("city" = province_state) %>% 
  filter (country_region == "China", city == c("Shanghai", "Beijing", "Chongqing")) %>% 
  mutate(date = as.Date(date, format = "%m/%d/%y"))

china_deaths %>% 
  ggplot(aes(x = deaths, fill = city)) + 
  geom_histogram(position = "dodge", binwidth = 2) +
  labs(title = "Distribution of Deaths from COVID-19 Between 1/22/20-3/7/20",
       x = "Number of Deaths",
       y = "Count")
```


### While a histogram is a great way of showing the distribution of data, having the bars for each group side-by-side gets sort of hard to understand. We can often use density plots in place of histograms when comparing distributions across groups:
```{r densityplot_china}
china_deaths %>% 
  ggplot(aes(x = deaths, fill = city)) + 
  geom_density(alpha = 0.4, adjust = 0.5, color = "black") +
  labs(title = "Distribution of Deaths from COVID-19 Between 1/22/20-3/7/20",
       x = "Number of Deaths",
       y = "Count")
```


### Your turn: Create both a histogram and a density plot of Italy's deaths from COVID-19 to date.
```{r deathdistribution_italy}

```



## Homework 1, Part 1: Finish generating the plots from the labs if you have not done so already. Knit this R Markdown document to a Word document and submit on Courseworks.